Ensemble Super-Learning for State Estimation. \\Application to the reconstruction of urban pollution.

### 2023-09-11 paper/main.tex
First final version of the paper.

As a guiding example, we consider the reconstruction of pollution maps in the city of Paris. \asmodif{We use the metric graph $\G$ provided by Open Street Maps and work with the Python library osmnx. We have worked with the graph covering the whole inner ring of the city as shown in Figure \ref{fig:paris-stations}. The full graph has $n(\V)=\perplexityinsert{numnodes}$ vertices and $n(\E)=\perplexityinsert{numedges}$ edges but we work with the biggest connected component of the subgraph that remains after filtering out all the edges whose associated streets have never been colored with traffic iformation. After this operation the actual graph used has $n(\V)=\perplexityinsert{numnodesfilter}$ vertices and $n(\E)=\perplexityinsert{numedgesfilter}$ edges}{}. The street network is relatively dense: the node degree histogram of Figure \ref{fig:degree-histogram} reveals that most nodes are connected to 3 to 6 edges. The vertices $\vv \in \V$ come with precise geographical coordinates. In the following, we assume that the graph is embedded in a two dimensional plane ($d=2$), and we do not take altitude into account. Each edge $\ee\in \E$ is a street or a portion of it, and we have access to its length $\ell_\ee$ and details about the shape of the street path. The information is so detailed that the streets are represented by paths that are not necessarily straight lines. However, in the following, some of our operations have assumed straight lines so we will work under this slightly simplified setting.

@@

We have worked with $m=\perplexityinsert{numstations}$ sensors functioning in the region of interest. As already brought up, their locations do not exactly match with the topology of the metric graph, therefore we project the position to the nearest node on the graph in a pre-processing step. This yields the observational nodes $\vv^\obs$ and coordinates $r^\obs$. As Figure \ref{fig:paris-stations} illustrates, the projected locations are very close to the exact locations given the global scale of the graph \asmodif{(the maximal discrepancy is $\perplexityinsert{MinDistNodeStation}$m, see \cref{fig:correlation})}{}. %Note however that the stations are originally located at different altitudes and since our graph is embedded only in $\bR^2$, we are not taking the effects of altitude into account.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Pre-processing of traffic data}
\label{sec:traffic-preprocess}
\asmodif{
We work with traffic information extracted from Google Maps images. On the one hand, the information is very rich because it is available in real time, and for the whole city (see Figure \ref{fig:traffic-raw-data}). On the other hand, the data is only partial: it comes in the form of four colors, and each color represents a certain traffic intensity. Therefore we only get an estimation about how many cars are currently on the street. In addition, the estimates are relative in the sense that the exact mapping between the colors and the number of cars are unknown. We only know that, e.g., red is more congested than green. We use this information to build estimates about the online traffic emissions by implementing the following pipeline:
}{}
\begin{itemize}
      \item \textbf{Raw data:} \asmodif{we made a script using the Selenium python library to automatically take screenshots of Paris\footnote{We tried in a first attempt to use the Google Maps Platform but as it required payment after one week of use we decided to code our own pipeline.} every $\perplexityinsert{screenshotperiod}$ minutes over an area of $\perplexityinsert{screenshotwidth} \times \perplexityinsert{screenshotheight}$ pixels with zoom $\perplexityinsert{zoom}$. An example of the raw  \textit{image} can be seen in \cref{fig:traffic-raw-data}. Note that city landmarks could not be removed before taking the screenshot neither by substracting a background image as each screenshot has slight variations rendering this approach impractical. Another issue is the absence of traffic data in the inner streets of the city. 
      }{}
      \item \textbf{Crop and extract colors:} \asmodif{for each image we first crop it to shape $\perplexityinsert{imagex} \times \perplexityinsert{imagey}$ to eliminate elements in the border that appear in the raw image like toolbars, minimap, etc. As the four colors associated with the four levels of traffic ($\{ \texttt{green},\, \texttt{orange},\, \texttt{red},\, \texttt{dark-red} \}$ \footnote{The RGB values of each color are: $\texttt{green}=\perplexityinsert{green}$, $\texttt{orange}=\perplexityinsert{yellow}$, $\texttt{red}=\perplexityinsert{red}$, $\texttt{dark-red}=\perplexityinsert{darkred}$.} in increasing order of traffic intensity) are clearly separated from the rest of the colors in a screenshot, we can track over images and extract the pixels that at some point had been colored with traffic color. This pixel locations almost perfectly overlap with the Open Street Maps' metric graph. The task is then to project the information on the graph.
}{}
      \item \textbf{Projection on graph:} For each image pixel, we can derive its geographical coordinate location, and find the closest edge from the street graph and consequently project the colored pixels presenting traffic to the closest edge. This way, we can build a function $q:\G\to \bR$ in the metric graph about the traffic level. For every edge $\ee\in \E$, the information is so detailed that the local traffic $r_\ee \to q_{|\ee}$ can be defined as a function varying with the coordinate $r_\ee$ of the edge. For the sake of simplicity, we work with constant values per edge: for each edge $\ee\in \E$, and each color $\texttt{c}\in \texttt{colors}$, \asmodif{we count the number of pixels with the corresponding color
      $$
      q_\ee(\texttt{c}) = \text{Number of pixels projected to $\ee$ with color $\texttt{c}$}.
      $$
      }{}
      %$$
      %r_\ee(\texttt{c}) = \frac{\text{Number of pixels projected to $\ee$ with color $\texttt{c}$}}{\text{Number of pixels projected to $\ee$}}
      %\; \in [0, 1].
      %$$
      %We next fix positive weights $w_{\texttt{c}}\geq 0$ such that
      %$$
      %\sum_{\texttt{c} \in \texttt{colors}} w_{\texttt{c}} = 1,
      %$$
      %\om{[Perhaps the weights do not sum up to one if we build a mapping to emissions, no?]} and compute the weighted average
      %$$
      %q_\ee = \sum_{\texttt{c} \in \texttt{colors}} w_{\texttt{c}} r_\ee(\texttt{c})
      %$$
      In this way, for each color, we have a piecewise constant function on the edges of the form
      \begin{equation}
      q(r,c)=\sum_{\ee\in \E} q_\ee(c) \mathds{1}_\ee(r).
      \label{eq:q}
      \end{equation}
      %An illustration of $q$ given weights $w_\texttt{c}$ is given in Figure \ref{fig:traffic-cars-on-edge}.
      \item \textbf{Traffic intensity-to-emissions mapping:} \asmodif{To transform the above data into a meaningful physical quantity, like and estimated average of cars emissions, one has to choose for this mapping a model that can be fixed beforehand, using expert knowledge, or let it be inferred directly from data which can be achieved by learning a simple linear model or a non linear one like a neural network}{}.

@@

Given the above definitions, the best reconstruction mapping is defined as
\begin{equation}
\label{eq:best-A}
\min_{A: \cX \to V} e_\star(A), \quad \star = \{\infty, \av\}
\end{equation}
Unfortunately, in practice it is not possible to find the best $A$ because:
\begin{itemize}
\item The set $\cM$ is unknown. Our only information comes from $X_T$ which is connected to a finite set of configurations $u_1, \dots, u_T$ from $\cM$, and it also contains partial information.
\item For every $u\in \cM$, we cannot evaluate the error $e(u, A)$ given that we do not have access to $u$ (so we cannot evaluate $u(r)$ for all $r\in \G$ as the norm $\Vert u - A(x_u)\Vert$ would require).
\item  The above points make it impossible to compute the wort-case error $e_\infty(A)$. The average error $e_\av(A)$ cannot be computed either because, in addition to all the previous points, $\pi$ is also unknown.
\end{itemize}
\asmodif{
Since computing the best $A$ in the sense of \eqref{eq:best-A} is out of reach, we inevitably have to build sub-optimal mappings as it is not possible to compare their performance rigorously according to our above quality criteria. Due to this, we propose to use nested $n$-fold cross-validation techniques which will be explained in the following section.
}{}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Validation set-up}
To validate the results we can only rely on the few available sensors ($m=\perplexityinsert{numstations}$) and it does not suffice to make a split of training and testing in time because we observed that models tend to over-fit the locations in which they where trained. For this we need to perform a tripple Leave-One-Out procedure on the stations: one for testing, one for training super-learners and the final one for trainig each individual model (see \cref{fig:loodiagram} for a schematic view of the successive splitings).
\begin{enumerate}
	\item \textbf{Test split}: The first split is performed both in time and space. We take the first $\perplexityinsert{percentagetraintime}\%$ of the available times ($\perplexityinsert{numberoftraintimes}$ hours) for training $T_{\train}$ and the rest for testing $T_{\test}$. The Leave-One-Out is done on the station locations $r^\obs$ and associated pollution measurements $z$. As some stations are too close to the border of the image, the available traffic information is not enough to add meaningfull information. For this reason even though we use all stations as information, we only take $\perplexityinsert{numstationstest}$ for perfoming the testing Leave-One-Out\footnote{Stations away from borders are: \perplexityinsert{stationstest}.}. This means that we have to train in total $\perplexityinsert{numstationstest}$ different sets of models associated with the different testing data splits $s\in S$, $\#S=\perplexityinsert{numstationstest}$, $I_s=\{i \in_{i \neq s} \{1, \dots m\}\}$ \om{[The idea is good. Something in the notation is maybe off. Some notation below was not introduced.]}
\begin{align}
	Y^{\test}_s(t) &= (z_s(t), r^\obs_s) \in \bR \times \bR^{2} \, \forall t \in T_{\test} \\
	X^{\train}_s(t) &= (z_{I_s}(t), r^\obs_{I_s}, q, T, w) \in \bR^{m-1} \times \bR^{2(m-1)} \times V \times \bR\times \bR^2 \, \forall t \in T_{\train}.	
\end{align}
	So the test error will be calculated as follws:
	$$e(u,A) \approx e^{\test}_{LOO}(Y^{\test}_s, A) = \frac{1}{\#T_{\test}}\sum_{t\in T_{\test}}\left(z_s-A(X^{\train}_s)(r^\obs_s)\right)^2$$
    \item \textbf{Super-learner train split}: For each training set we perform another Leave-One-Out to train each super-learner $(Y^{SL}_s, X^{SL}_s)$ both in $X^{\train}_s$.
    \item \textbf{Individual models train split}: For each training super-learner set we perform eone more Leave-One-Out to train each individual model under different circumstances avoiding this way the possibility of over-fitting: $(Y^{M}_s, X^{M}_s)$ both in $X^{SL}_s$.
\end{enumerate}

\ascomment{[Systematic way of writing this successive subdivisions and minimization errors.]}

@@

The most usual approach to find the parameters is to resort to cross-validation, in which we search for
\ascomment{Check the equations, is it well written formally the belonging to the different LOO splits?} \om{[Which LOO do we use here?]}
\begin{equation}
\label{eq:fit-linear}
\min_{ \substack{\theta_1,\dots, \theta_n \in \Theta \\ \gamma_1,\dots, \gamma_n \in \Gamma}}
\sum_{ \substack{x\in X^{M} \\ (z, r)\in Y^{M}}}
\Vert z - \sum_{i=1}^n \ell(x, \theta_i) v(r, \gamma_i) \Vert^2_2,
\end{equation}
where $\Vert \cdot \Vert_2$ denotes the Euclidean norm. This norm can of course be replaced by any other lower semi-continuous loss function. From the optimal parameters $\theta_i^*$ and $\gamma_i^*$ of \eqref{eq:fit-linear}, we define
$$
A(x)(r) = \sum_{i=1}^n \ell(x, \theta_i^*) v(r, \gamma^*_i).
$$
Note that when using reduced modeling techniques, there is no optimization over the $\gamma_i$ since the $v_i$ are selected beforehand with specific techniques. A classical linear reconstruction map which we consider in our work is the so-called PBDW method (Parametrized-Background Data-Weak approach, see \cite{MPPY2015, BCDDPW2017, Mula2023}).

\paragraph{Nonlinear mappings:} In this case, $A$ is searched in a family of parametrized mappings of the form
$$
\cF \coloneqq \{ D(\theta): \cX \to V   \cond \theta \in \Theta  \},
$$
where the decoder function $D:\cX\to V$ is not necessarily linear. A common choice is to take $D$ as neural-network $\cN\cN$ with a given architecture, and trainable parameters $\theta \in \Theta\subseteq \bR^n$. By computing
\ascomment{Check the equations, is it well written formally the belonging to the different LOO splits?} \om{[What LOO to use?]}
$$
\theta^* \in \argmin_{\theta \in \Theta}
\sum_{\substack{x\in X^{M} \\ (z, r)\in Y^{M}}}
\Vert z - \cN\cN(\theta)(r, x) \Vert^2_2,
$$
we then set
$$
A = \cN\cN(\theta^*).
$$
Note that the input and output spaces of the neural network are not finite dimensional spaces like in the classical definition of neural networks. Instead, our neural network should be understood as a neural operator. In section \ref{sec:concrete-methods} we provide a concrete example for our application.

Another option as a nonlinear map are nonlinear extensions of the PBDW approach (see \cite{CDMN2022,CDMS2022}).

@@

Similarly to Section \ref{sec:mappings}, we can again build linear and nonlinear super-learners. The most simple possibility is to consider a linear combination (usually called aggregation in statistics),
$$
\cS(A_1, \dots, A_p) = \sum_{i=1}^p \sigma_i A_i,
$$
and we then search for the best coefficients $\sigma_1, \dots, \sigma_p$ with a leave-one-out procedure. We can alternatively introduce a non linear mapping (for example, a neural network) that also depends on parameters $\sigma$ to be learnt and that transforms the individual learners outputs into a single prediction. Defining the vector
$$
c(x, r) = ( A_1(x)(r), \dots, A_p(x)(r) )^T \in \bR^p, \quad \forall (x,r)\in \cX\times \G,
$$
we define
$$
\cS(A_1, \dots, A_p)(x)(r) =  \mathcal{NN} (\sigma^*)( c(x,r) ). 
$$
where $\sigma^*$ is found by an $n$-fold leave-one out,
\asmodif{
$$
\sigma^* \in \underset{\sigma \in \Sigma}{\argmin}
\sum_{\substack{x\in X^{SL} \\ (z, r)\in Y^{SL}}}
\Vert z -  \mathcal{NN} (\sigma^*)( c(x,r) ) \Vert^2_2.
$$
Note that here we use the first training set $X^{SL}$ instead of the second one which is used to train each learner.}{} This is done in order to avoid over-fitting.

@@

\asmodif{
Finally, another way to build a super-learner is by sequentially composing individual models $A_j$ together with transition functions $P_j$ and train them to correct the errors of previous ones so that \om{[It is $A_j(x)$ and not $A_j(\theta)$. I guess you mean to discuss the case when $A_j = \cN\cN(\theta)?$]}
$$
S_j(A_1, \dots, A_j) = P_j(\sigma_j^*) \circ A_j(\theta_j^*) \circ \dots \circ P_1(\sigma_1^*) \circ A_1(\theta_1^*)
$$
and $S=S_p$. At each step $j$ the target of each submodel is to correct the residuals $\tilde z_j=z - \cS_{j-1}(r, x)$ of previous steps, so $(z, r)\in Y^{M} \mapsto (\tilde z_j^{Y^{M}}, r) \in \tilde Y^{M}$. At the same time the information received from the pollution part of the input $x$ is also replaced by the last residual $x = (z, r^\obs, q, T, w) \in X^{M} \mapsto x_j^M = (\tilde z_j^{X^{M}}, r^\obs, q, T, w) \in \tilde X^{M}$ so the minimization problem in this step becomes
$$
\theta_j^* \in \underset{\theta_j \in \Theta_j}{\argmin}
\sum_{x_j^M, \tilde z_j^{Y^{M}}, r}
\Vert \tilde z_j^{Y^{M}} - A_j(\theta_j)(r, x_j^M)) \Vert^2_2.
$$
The transition maps $P_j$ are also trained to correct the residuals $\hat z_j=z - A_j(\theta_j^*)\circ\cS_{j-1}(r, x)$ of previous steps, so $(z, r)\in Y^{SL} \mapsto (\hat z_j^{Y^{SL}}, r) \in \hat Y^{SL}$ and $x = (z, r^\obs, q, T, w) \in X^{SL} \mapsto x_j^{SL} = (\tilde z_j^{X^{SL}}, r^\obs, q, T, w)  \in \hat X^{SL}$ so the minimization problem in this step becomes
$$
\sigma_j^* \in \underset{\sigma_j \in \Sigma_j}{\argmin}
\sum_{\hat z_j^{X^{SL}}, \hat z_j^{Y^{SL}}, r}
\Vert \hat z_j^{Y^{SL}} - P_j(\sigma_j)(\hat z_j^{X^{SL}}) \Vert^2_2.
$$
}{}

@@

As the traffic information $q$ is supported on the edges of the street graph of the city of Paris one could extract meaningful information about the vicinity of some target point $x$ by doing some kind of convolution on the graph. This leads us naturally to the idea of Graph Convolution by which one can do meaningfull feature extraction for our problem.

@@

\begin{figure}
    \centering
    \includegraphics[width=0.9\linewidth]{figures/plot_pollution_map_in_graph_modelLR_Extra_stationOPERA.pdf}
    \includegraphics[width=0.9\linewidth]{figures/plot_pollution_map_in_graph_modelNN_Extra_stationOPERA.pdf}
    \caption{Pollution maps.}
    \label{fig:pollutionmaps}
\end{figure}

>>>>

As a guiding example, we consider the reconstruction of pollution maps in the city of Paris. We use the metric graph $\G$ provided by Open Street Maps and work with the Python library osmnx. We have worked with the graph covering the whole inner ring of the city as shown in Figure \ref{fig:paris-stations}. The full graph has $n(\V)=\perplexityinsert{numnodes}$ vertices and $n(\E)=\perplexityinsert{numedges}$ edges but we work with the biggest connected component of the subgraph that remains after filtering out all the edges whose associated streets have never been colored with traffic iformation. After this operation the actual graph used has $n(\V)=\perplexityinsert{numnodesfilter}$ vertices and $n(\E)=\perplexityinsert{numedgesfilter}$ edges. The street network is relatively dense: the node degree histogram of Figure \ref{fig:degree-histogram} reveals that most nodes are connected to 3 to 6 edges. The vertices $\vv \in \V$ come with precise geographical coordinates. In the following, we assume that the graph is embedded in a two dimensional plane ($d=2$), and we do not take altitude into account. Each edge $\ee\in \E$ is a street or a portion of it, and we have access to its length $\ell_\ee$ and details about the shape of the street path. The information is so detailed that the streets are represented by paths that are not necessarily straight lines. However, in the following, some of our operations have assumed straight lines so we will work under this slightly simplified setting.

@@

We have worked with $m=\perplexityinsert{numstations}$ sensors functioning in the region of interest. As already brought up, their locations do not exactly match with the topology of the metric graph, therefore we project the position to the nearest node on the graph in a pre-processing step. This yields the observational nodes $\vv^\obs$ and coordinates $r^\obs$. As Figure \ref{fig:paris-stations} illustrates, the projected locations are very close to the exact locations given the global scale of the graph (the maximal discrepancy is $\perplexityinsert{MinDistNodeStation}$m, see \cref{fig:correlation}). %Note however that the stations are originally located at different altitudes and since our graph is embedded only in $\bR^2$, we are not taking the effects of altitude into account.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Pre-processing of traffic data}
\label{sec:traffic-preprocess}

We work with traffic information extracted from Google Maps images. On the one hand, the information is very rich because it is available in real time, and for the whole city (see Figure \ref{fig:traffic-raw-data}). On the other hand, the data is only partial: it comes in the form of four colors, and each color represents a certain traffic intensity. Therefore we only get an estimation about how many cars are currently on the street. In addition, the estimates are relative in the sense that the exact mapping between the colors and the number of cars are unknown. We only know that, e.g., red is more congested than green. We use this information to build estimates about the online traffic emissions by implementing the following pipeline:
\begin{itemize}
      \item \textbf{Raw data:} we made a script using the Selenium python library to automatically take screenshots of Paris\footnote{We tried in a first attempt to use the Google Maps Platform but as it required payment after one week of use we decided to code our own pipeline.} every $\perplexityinsert{screenshotperiod}$ minutes over an area of $\perplexityinsert{screenshotwidth} \times \perplexityinsert{screenshotheight}$ pixels with zoom $\perplexityinsert{zoom}$. An example of the raw  \textit{image} can be seen in \cref{fig:traffic-raw-data}. Note that city landmarks could not be removed before taking the screenshot neither by substracting a background image as each screenshot has slight variations rendering this approach impractical. Another issue is the absence of traffic data in the inner streets of the city.
      \item \textbf{Crop and extract colors:} for each image we first crop it to shape $\perplexityinsert{imagex} \times \perplexityinsert{imagey}$ to eliminate elements in the border that appear in the raw image like toolbars, minimap, etc. As the four colors associated with the four levels of traffic ($\{ \texttt{green},\, \texttt{orange},\, \texttt{red},\, \texttt{dark-red} \}$ \footnote{The RGB values of each color are: $\texttt{green}=\perplexityinsert{green}$, $\texttt{orange}=\perplexityinsert{yellow}$, $\texttt{red}=\perplexityinsert{red}$, $\texttt{dark-red}=\perplexityinsert{darkred}$.} in increasing order of traffic intensity) are clearly separated from the rest of the colors in a screenshot, we can track over images and extract the pixels that at some point had been colored with traffic color. This pixel locations almost perfectly overlap with the Open Street Maps' metric graph. The task is then to project the information on the graph.

      \item \textbf{Projection on graph:} For each image pixel, we can derive its geographical coordinate location, and find the closest edge from the street graph and consequently project the colored pixels presenting traffic to the closest edge. This way, we can build a function $q:\G\to \bR$ in the metric graph about the traffic level. For every edge $\ee\in \E$, the information is so detailed that the local traffic $r_\ee \to q_{|\ee}$ can be defined as a function varying with the coordinate $r_\ee$ of the edge. For the sake of simplicity, we work with constant values per edge: for each edge $\ee\in \E$, and each color $\texttt{c}\in \texttt{colors}$, we count the number of pixels with the corresponding color
      $$
      q_\ee(\texttt{c}) = \text{Number of pixels projected to $\ee$ with color $\texttt{c}$}.
      $$
      %$$
      %r_\ee(\texttt{c}) = \frac{\text{Number of pixels projected to $\ee$ with color $\texttt{c}$}}{\text{Number of pixels projected to $\ee$}}
      %\; \in [0, 1].
      %$$
      %We next fix positive weights $w_{\texttt{c}}\geq 0$ such that
      %$$
      %\sum_{\texttt{c} \in \texttt{colors}} w_{\texttt{c}} = 1,
      %$$
      %\om{[Perhaps the weights do not sum up to one if we build a mapping to emissions, no?]} and compute the weighted average
      %$$
      %q_\ee = \sum_{\texttt{c} \in \texttt{colors}} w_{\texttt{c}} r_\ee(\texttt{c})
      %$$
      In this way, for each color, we have a piecewise constant function on the edges of the form
      \begin{equation}
      q(r,c)=\sum_{\ee\in \E} q_\ee(c) \mathds{1}_\ee(r).
      \label{eq:q}
      \end{equation}
      %An illustration of $q$ given weights $w_\texttt{c}$ is given in Figure \ref{fig:traffic-cars-on-edge}.
      \item \textbf{Traffic intensity-to-emissions mapping:} To transform the above data into a meaningful physical quantity, like and estimated average of cars emissions, one has to choose for this mapping a model that can be fixed beforehand, using expert knowledge, or let it be inferred directly from data which can be achieved by learning a simple linear model or a non linear one like a neural network.

@@

Given the above definitions, the best reconstruction mapping is defined as
\begin{equation}
\label{eq:best-A}
\min_{A: \cX \to V} e_\star(A), \quad \star = \{\infty, \av\}
\end{equation}
Unfortunately, in practice it is not possible to find the best $A$ because:
\begin{itemize}
\item The set $\cM$ is unknown. Our only information comes from $X_T$ which is connected to a finite set of configurations $u_1, \dots, u_T$ from $\cM$, and it also contains partial information.
\item For every $u\in \cM$, we cannot evaluate the error $e(u, A)$ given that we do not have access to $u$ (so we cannot evaluate $u(r)$ for all $r\in \G$ as the norm $\Vert u - A(x_u)\Vert$ would require).
\item  The above points make it impossible to compute the wort-case error $e_\infty(A)$. The average error $e_\av(A)$ cannot be computed either because, in addition to all the previous points, $\pi$ is also unknown.
\end{itemize}
Since computing the best $A$ in the sense of \eqref{eq:best-A} is out of reach, we inevitably have to build sub-optimal mappings as it is not possible to compare their performance rigorously according to our above quality criteria. Due to this, we propose to use nested $n$-fold cross-validation techniques which will be explained in the following section.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Validation set-up}
To validate the results we can only rely on the few available sensors ($m=\perplexityinsert{numstations}$) and it does not suffice to make a split of training and testing in time because we observed that models tend to over-fit the locations in which they where trained. For this we need to perform a triple Leave-One-Out procedure on the stations: one for testing, one for training super-learners and the final one for training each individual model (see \cref{fig:loodiagram} for a schematic view of the successive splittings).

% \ascomment{[Attempt to systematize and formalize the iterative-LOO, the successive subdivisions and minimization errors.]}

% Split $\cD$: starting from space $X_j^H \subseteq X$ \ascomment{It is not exactly a subset but a concatenation of sub-partitions} with $H$ the set of available times in $X_j^H$  we will construct space $\cD(X_j^H)=X_{j+1}^H\times Y_{j+1}^H$ as the concatenation of its 1-knock-out-element partitions $X_{j+1}^H\times Y_{j+1}^H=^\frown_{s\in S} X_{js}^H \times Y_{js}^H=^\frown_{s\in S}\cD_s(X_j^H)$ \ascomment{which symbol do we use for concatenation?} \ascomment{It is not a tensor product $\times$ but an "alignement", maybe better $(X_{j+1}^H, Y_{j+1}^H)$} where $s\in S_j$ with $S_j$ being the set of available stations in $X_j^H$, $I_s=\{i \in_{i \neq s} S\}$ and
% \begin{align}
%     Y_{js}^H &= (z_s(t), r^\obs_s) \in \bR \times \bR^{2} \, \forall t \in H\\
%     X_{js} &= (z_{I_s}(t), r^\obs_{I_s}, q, T, w) \in \bR^{\#S_j-1} \times \bR^{2(\#S_j-1)} \times V \times \bR\times \bR^2 \, \forall t \in H
% \end{align}
% Then
% \begin{align}
% X^{\test}_s\times Y^{\test}_s &= \cD_s(X^{T_\test})\\
% X^{\train}\times Y^{\train} &= \cD(X^{T_\train})\\
% X^{SL}\times Y^{SL} &= \cD(X^{\train})\\
% X^{M}\times Y^{M} &= \cD(X^{SL})\\
% \end{align}

\begin{enumerate}
	\item \textbf{Test split}: The first split is performed both in time and space. We take the first $\perplexityinsert{percentagetraintime}\%$ of the available times ($\perplexityinsert{numberoftraintimes}$ hours) for training $T_{\train}$ and the rest for testing $T_{\test}$. The Leave-One-Out is done on the station locations $r^\obs$ and associated pollution measurements $z$. As some stations are too close to the border of the image, the available traffic information is not enough to add meaningfull information. For this reason even though we use all stations as information, we only take $\perplexityinsert{numstationstest}$ for perfoming the testing Leave-One-Out\footnote{Stations away from borders are: \perplexityinsert{stationstest}.}. This means that we have to train in total $\perplexityinsert{numstationstest}$ different sets of models each one associated with one of these different testing data splits. Let $m_\test = \perplexityinsert{numstationstest}$, and let
    $$
    I = \{1, \dots, m_\test\}
    $$
    and let
    $$
    [I]^k = \{ S\subseteq I \cond \#S=k \}
    $$
    be the set of all $k$-subsets of $I$ (for $1\leq k \leq m_\test$). Note that we have $\#[I]^k = \binom{m_\test}{k}$ subsets in $I$ and in particular $k=1$ for a Leave-one-out. For any $S\subseteq [I]^k$,
    \begin{align}
    X^{\train}_S(t)
    &= (z_{S^c}(t), r^\obs_{S^c}, q, T, w) \, \quad\forall t \in T_{\train}. \\
    % X^{\test}_S(t)
    % &= (z_{S}(t), r^\obs_{S}, q, T, w) \, \quad\forall t \in T_{\test}. \\
    Y^{\test}_S(t)
    &= (z_{S}(t), r^\obs_{S}) \in \bR \times \bR^{2} \, \forall t \in T_{\test}
    \end{align}
    $$
    e_S(u)
    \approx
    \frac{1}{T_\test %\binom{\perplexityinsert{numstationstest}}{k}
    }
    \sum_{t\in T_\test}
    \Vert z_{S}(t) - A(X_{S^c}^\train)(r_{S}^\obs)\Vert^2
    $$
    \item \textbf{Super-learner train split}: Similarly, for each training set we perform another Leave-One-Out but now we train each super-learner on the concatenation of all the $m_\test-1$ partitions: $(Y^{SL}_{S}, X^{SL}_{S}) = \underset{S'\subset S^c \, : \, \#S'=1}{^\frown}(Y^{SL}_{SS'}, X^{SL}_{SS'})$ both in $X^{\train}_S$.
    \item \textbf{Individual models train split}: For each training super-learner set $(Y^{SL}_{S}, X^{SL}_{S})$, both in $X^{SL}_S$, we perform in a similar way one more Leave-One-Out to train each individual model to avoid the possibility of over-fitting.
\end{enumerate}

@@

The most usual approach to find the parameters is to resort to cross-validation, in which we search for

\begin{equation}
\label{eq:fit-linear}
\min_{ \substack{\theta_1,\dots, \theta_n \in \Theta \\ \gamma_1,\dots, \gamma_n \in \Gamma}}
\sum_{ \substack{\left(x,(z, r)\right)\in (X^{M}, Y^{M})}}
\Vert z - \sum_{i=1}^n \ell(x, \theta_i) v(r, \gamma_i) \Vert^2_2,
\end{equation}
where $\Vert \cdot \Vert_2$ denotes the Euclidean norm. This norm can of course be replaced by any other lower semi-continuous loss function. From the optimal parameters $\theta_i^*$ and $\gamma_i^*$ of \eqref{eq:fit-linear}, we define
$$
A(x)(r) = \sum_{i=1}^n \ell(x, \theta_i^*) v(r, \gamma^*_i).
$$
Note that when using reduced modeling techniques, there is no optimization over the $\gamma_i$ since the $v_i$ are selected beforehand with specific techniques. A classical linear reconstruction map which we consider in our work is the so-called PBDW method (Parametrized-Background Data-Weak approach, see \cite{MPPY2015, BCDDPW2017, Mula2023}).

\paragraph{Nonlinear mappings:} In this case, $A$ is searched in a family of parametrized mappings of the form
$$
\cF \coloneqq \{ D(\theta): \cX \to V   \cond \theta \in \Theta  \},
$$
where the decoder function $D:\cX\to V$ is not necessarily linear. A common choice is to take $D$ as neural-network $\cN\cN$ with a given architecture, and trainable parameters $\theta \in \Theta\subseteq \bR^n$. By computing
$$
\theta^*=\underset{\theta \in \Theta}{\argmin}
\sum_{\substack{\left(x,(z, r)\right)\in (X^{M}, Y^{M})}}
\Vert z - D(\theta)(r, x) \Vert^2_2,
$$
we then set
$$
A = D(\theta^*).
$$
Note that the input and output spaces of the neural network are not finite dimensional spaces like in the classical definition of neural networks. Instead, our neural network should be understood as a neural operator. In section \ref{sec:concrete-methods} we provide a concrete example for our application.

Another option as a nonlinear map are nonlinear extensions of the PBDW approach (see \cite{CDMN2022,CDMS2023}).

@@

Similarly to Section \ref{sec:mappings}, we can again build linear and nonlinear super-learners. The most simple possibility is to consider a linear combination (usually called aggregation in statistics),
$$
\cS(A_1, \dots, A_p) = \sum_{i=1}^p \sigma_i A_i,
$$
and we then search for the best coefficients $\sigma_1, \dots, \sigma_p$ with a leave-one-out procedure. We can alternatively introduce a non linear mapping (for example, a neural network) that also depends on parameters $\sigma$ to be learnt and that transforms the individual learners outputs into a single prediction. Defining the vector
$$
c(x, r) = ( A_1(x)(r), \dots, A_p(x)(r) )^T \in \bR^p, \quad \forall (x,r)\in \cX\times \G,
$$
we define
$$
\cS(A_1, \dots, A_p)(x)(r) =  \mathcal{NN} (\sigma^*)( c(x,r) ).
$$
where $\sigma^*$ is found by a leave-one out procedure,

$$
\sigma^* = \underset{\sigma \in \Sigma}{\argmin}
\sum_{\substack{(x, (z, r))\in (X^{SL}, Y^{SL})}}
\Vert z -  \cS (\sigma^*)( c(x,r) ) \Vert^2_2.
$$
Note that here we use the first training set $X^{SL}$ instead of the second one which is used to train each learner. This is done in order to avoid over-fitting.

@@

Finally, another way to build a super-learner is by sequentially composing individual models $A_j$ together with transition functions $P_j$ and train them to correct the errors of previous ones so that
$$
S_j(A_1, \dots, A_j) = P_j(\sigma_j^*) \circ A_j \circ \dots \circ P_1(\sigma_1^*) \circ A_1
$$
and $S=S_p$. At each step $j$ the target of each submodel is to correct the residuals $\tilde z_j=z - \cS_{j-1}(r, x)$ of previous steps, so $(z, r)\in Y^{M} \mapsto (\tilde z_j^{Y^{M}}, r) \in \tilde Y^{M}$. At the same time the information received from the pollution part of the input $x$ is also replaced by the last residual $x = (z, r^\obs, q, T, w) \in X^{M} \mapsto x_j^M = (\tilde z_j^{X^{M}}, r^\obs, q, T, w) \in \tilde X^{M}$ so the minimization problem in this step becomes
$$
\theta_j^* = \underset{\theta_j \in \Theta_j}{\argmin}
\sum_{x_j^M, \tilde z_j^{Y^{M}}, r}
\Vert \tilde z_j^{Y^{M}} - D(\theta_j)(r, x_j^M)) \Vert^2_2.
$$
and then $A_j=D(\theta_j^*)$.

The transition maps $P_j$ are also trained to correct the residuals $\hat z_j=z - D_j(\theta_j^*)\circ\cS_{j-1}(r, x)$ of previous steps, so $(z, r)\in Y^{SL} \mapsto (\hat z_j^{Y^{SL}}, r) \in \hat Y^{SL}$ and $x = (z, r^\obs, q, T, w) \in X^{SL} \mapsto x_j^{SL} = (\tilde z_j^{X^{SL}}, r^\obs, q, T, w)  \in \hat X^{SL}$ so the minimization problem in this step becomes
$$
\sigma_j^* = \underset{\sigma_j \in \Sigma_j}{\argmin}
\sum_{\hat z_j^{X^{SL}}, \hat z_j^{Y^{SL}}, r}
\Vert \hat z_j^{Y^{SL}} - P_j(\sigma_j)(\hat z_j^{X^{SL}}) \Vert^2_2.
$$

@@

As the traffic information $q$ is supported on the edges of the street graph of the city of Paris one could extract meaningful information about the vicinity of some target point $x$ by doing some kind of convolution on the graph. This leads us naturally to the idea of Graph Convolution by which one can do meaningful feature extraction for our problem.

@@

\begin{figure}
    \centering
    \includegraphics[width=0.9\linewidth]{figures/plot_pollution_map_in_graph_modelLR_Extra_stationOPERA.pdf}
    \includegraphics[width=0.9\linewidth]{figures/plot_pollution_map_in_graph_modelNN_Extra_stationOPERA.pdf}
    \caption{Pollution maps. \ascomment{Change color map to autumn}}
    \label{fig:pollutionmaps}
\end{figure}
